The Goldilocks Level of Friction: Using User Numbers, Ratings, and Reviews to Evaluate Digital Self-Control Tools

### 2022-04-29 paper.Rmd
use visual editor

In academia, a growing amount of HCI research is focusing on developing and evaluating various design patterns meant to help users self-regulate digital device use [@lottridgeBrowserDesignImpacts2012; @kimGoalKeeperExploringInteraction2019; @pinderPushAwaySmartphone2019; @tsengOvercomingDistractionsTransitions2019; @mongeroffarelloRaceDigitalWellbeing2019; @lyngsSelfControlCyberspaceApplying2019]^[We use the term 'design patterns' to refer to specific functions intended to support people in self-regulating their digital device use (e.g., hiding newsfeeds, providing goal reminders), and the term ‘digital self-control tool’ (DSCT) to refer to apps and browser extensions that implement a given design pattern (e.g., a browser extension for hiding Facebook’s newsfeed). We occasionally use the term 'intervention' to refer to the implementation of a design pattern in the context of it changing a user’s usual digital environment, or when it is evaluated in a research study.].
This area of HCI work has gradually been building an understanding of how a range of design patterns from visualisations of time spent [@whittakerDonWasteMy2016] to goal-setting with social support [@koNUGUGroupbasedIntervention2015] can help people stay in control.
However, we are still in the early days of answering basic questions including how to define and measure 'success' for these design patterns, how much people vary in how useful they find particular interventions, and how specific design patterns' usefulness depends on details of their implementation, such as the level of 'friction' they introduce [@kimGoalKeeperExploringInteraction2019; @markEffectsIndividualDifferences2018; @tseng2019; @lyngsTellMeWhat2018; @hinikerMovingGreatScreen2019].

@@

Users want DSCTs to flexibly adapt to their personal definition of ‘distraction’, as seen in common requests for customisation of *what* and *when* particular usage patterns should be blocked or tracked.
A key reason that users' and tools' conceptualisation of distraction should be in close alignment is that DSCTs' interventions, useful as they may be, can elicit negative emotional reactions. This may cause the effects of interventions to be counterproductive, especially when users perceive DSCTs' goals or framing to differ from their own.

Finally, users want DSCTs to provide carefully calibrated incentives that are sufficient to incentivise intended behaviour change without feeling overly intrusive.
In our analysis, this 'Goldilocks level'^[After the fairy tale 'Goldilocks and the Three Bears', where a young girl tastes three different bowls of porridge and prefers the one that is neither too hot nor too cold, but 'just right'. The concept of the 'just right amount' is often referred to as the 'Goldilocks principle'. For example, a widely cited paper on the relationship between screen time and mental well-being documented that a balance point in between "too much" and "too little" tech use seemed optimal, which they referred to as the 'digital Goldilocks hypothesis' [@Przybylski2017].] varied between reviewers, but they commonly wanted tools to provide incentives powerful enough to keep themselves on track in moments of temptation or absentmindedness.
Overly strong incentives could backfire if they simply made users feel like a failure (in the case of reward/punishment tools) or did not allow some way to handle genuine emergency use (in the case of block/removal tools).

@@

One approach to addressing these questions is to evaluate new or existing interventions in lab and field studies, and/or conduct surveys or interviews with users.
A complementary approach, which  we explore in the present paper, is to investigate the landscape of digital self-control tools available online.
Here, many hundreds of tools are exploring a wide range of strategies, from limiting the amount of functionality available on home screens (e.g.~the Android app \emph{LessPhone Launcher} [@mohanLessPhoneLauncher2019]) and hiding or modifying content on distracting websites (e.g. *Newsfeed Eradicator* [@jdevNewsFeedEradicator2019]), to gamifying self-control by tying device use to the wellbeing of virtual creatures (e.g. *Forest* [@seekrtechForestStayFocused2018]).

@@

We investigated the 367 digital self-control tools reviewed by Lyngs et al. [@lyngsSelfControlCyberspaceApplying2019], as this paper provided the largest currently available review of DSCTs.
Lyngs et al. identified these tools by downloading search results for the terms 'distraction', 'addiction', 'motivation', and 'self-control', individually as well as prefixed by 'internet', 'smartphone', and 'laptop', on the Chrome Web, Google Play, and Apple App stores.
From the initial results (4890 distinct apps and extensions), they included tools explicitly designed to help people self-regulate digital device use, resulting in a data set of 367 tools (222 browser extensions from Chrome Web, 87 apps from Google Play, and 58 apps from the Apple App store).^[See Lyngs et al. [@lyngsSelfControlCyberspaceApplying2019] and their supplementary materials on [osf.io/zyj4h/](https://osf.io/zyj4h/) for further details.]

@@

## Data and analysis approach
### User numbers and ratings
All stores displayed average ratings, as well as a count of how many ratings a tool had received.
For installation numbers, the stores provided different information: the Chrome Web store provided an exact number of users, and the Google Play store provided a 'minimum number of installs' (e.g. '100,000+').
The Apple App store provided no direct information about user numbers.
For processing which ranked tools by installation numbers, we used number of ratings as a proxy for relative numbers of users for tools on the Apple App store.

@@

- \typeBlock\hspace{1mm}= block/removal, \typeTrack\hspace{1mm}= self-tracking,\
\typeGoal\hspace{1mm}= goal-advancement, \typeReward\hspace{1mm}= reward/punish
- C = Chrome Web, P = Google Play, A = Apple App store
- Example review quote from a Chrome Web tool including design patterns for block/removal, self-tracking, and goal-advancement: *"It would be nice to make it hard to disable, something like a simple maths problem that we could change difficulty, so I am not tempted."*, R1408, Timewarp \typeBlock\typeTrack\typeGoal \textbf{C}

@@

An early observation as we explored reported benefits of use was that a substantial proportion of reviews expressed highly positive and important impact of the tools used (n = 480, 22% of reviews).
Thus, the reviews described a common feeling of lacking control over digital device use, which resulted in using devices in unintended ways, or in using them so much that it interfered with other goals.
However, digital self-control tools were seen as effectively helping reviewers take back control, save time, and improve their lives (*"I can happily say that this extension has changed my life!!!! It hides EVERY ANNOYING YouTube feature"*, R385, \href{https://chrome.google.com/webstore/detail/df-tube-distraction-free/mjdepdfccjgcndkmemponafgioodelna?hl=en}{DF Tube} \typeBlock \textbf{C}, *"Has saved me countless hours of time-wasting nonsense"*, R1000, NewsFeed Eradicator for Facebook \typeBlock\typeGoal \textbf{C}), and many expressed deep appreciation towards the developers ("Who so ever is working behind this software, my deepest gratitude to them", R271, \href{https://chrome.google.com/webstore/detail/blocksite-block-websites/eiimnmioipafcokbfikbljfdeojpcgbh?hl=en-GB}{\emph{BlockSite}} \typeBlock \textbf{C}). 
Some reviewers seemed surprised that particular tools had been useful to them (*"I added it more as a joke, but it's really good and helpful for those hours you're supposed to study and don't want to be distracted"*, R675, \href{https://chrome.google.com/webstore/detail/go-fucking-work/hibmkkpfegfiinilnlabbfnjcopdiiig}{Go Fucking Work} \typeBlock\typeReward \textbf{C}).
Importantly, the reviews also contained a wealth of information on how specific implementations of design patterns could fail to fulfil their purpose, and how they might be improved.

@@

Users saw DSCTs as critical to avoiding time-wasting patterns of habitual or impulsive digital device use.
Many reviews described this challenge as related to 'addictive' design patterns and other adversarial dynamics online, to which DSCTs provided essential countermeasures that reviewers had searched for (*"Was looking for a solution for this for over 1 year"*, R771, \href{https://chrome.google.com/webstore/detail/inbox-when-ready-for-gmai/cdedhgmbfjhobfnphaoihdfmnjidcpim}{Inbox When Ready for Gmail} \typeBlock\typeGoal\typeTrack \textbf{C}; *"Should be on all phones as standard"*, R982, Mute Screen Time Tracker \typeGoal\typeTrack \textbf{A}). 
DSCTs seemed especially helpful when it was not possible to avoid using distracting services (*"This helps me avoid the mind numbing news feed that is of Facebook when work has to be done on Facebook itself"*, R873, \href{https://chrome.google.com/webstore/detail/kill-news-feed/hjobfcedfgohjkaieocljfcppjbkglfd}{Kill News Feed} \typeBlock \textbf{C}).

The usage contexts mentioned suggested that, aside from wanting to limit time on devices overall or on specific services (n = 62, 4\% of reviews), reviewers used DSTCs to help them focus on a given task amidst all too easily available distractions on their devices (n = 201, 13\%).
The need for support from DSCTs to stay focused seemed especially acute when tasks were important, but less instantly rewarding, expressed in some reviews as DSCTs helping overcome 'procrastination' (*"seeing my short-term to-do list nudges me to get back to work when I am opening a new website to procrastinate"*, R324, Dayboard \typeBlock\typeTrack \textbf{C}). 
In these situations, DSCTs helped reviewers stay in control by reminding them of what they wanted to do by reducing exposure to 'temptations' (*"Without it, it is too easy for me to be tempted to view and open YouTube sidebar links"*, R376, DF Tube \typeBlock \textbf{C}), and/or by providing incentives for intended use (*"Not only does it help me stay focused, I smile with delight looking at all my trees and flowers at the end of the day"*, R642, Forest \typeTrack\typeReward \textbf{C}). 
Thus, many reviewers generically stated that DSCTs benefited them in productivity/work contexts (n = 167, 11\%; "Not being distracted by unread/misc emails has proven to be of significant benefit to my productivity", R776, \href{https://chrome.google.com/webstore/detail/inbox-when-ready-for-gmai/cdedhgmbfjhobfnphaoihdfmnjidcpim}{\emph{Inbox When Ready for Gmail}} \typeBlock\typeGoal\typeTrack \textbf{C}), including studying, writing, or reading (n = 90, 6\%).

Another important use context was \emph{sleep}, where many reviewers said their digital devices often interfered with going to bed at the intended time. However, DSCTs helped them adjust their habits in favour of the sleep patterns they wanted (*"I always used to stay up really late on my phone until I got this app - I have consistently been getting 9 hours sleep and waking up early for the first time in my life !!"*, R1252, [SleepTown](https://apps.apple.com/gb/app/sleeptown/id1210251567)\typeTrack\typeReward \textbf{P}; *"It really forces me to stay off to the phone, and so it's useful especially when I need to go to bed"*, R198, BlackOut \typeBlock \textbf{P}).

**Sub-theme: DSCTs as especially useful for people with more acute self-regulation challenges**
In an extension of the main theme, DSCTs seemed especially useful for individuals who experience self-regulation challenges more intensely.
This was often expressed as a tool being helpful for tackling 'addiction', and/or as reviewers labelling themselves 'addicts', 'procrastinators' or just low in self-control (n = 58, 3.8\%; *"This was the perfect solution to my phone addiction. It was consuming my life and my free time"*, R612, Forest \typeTrack\typeReward \textbf{P}).
Sometimes reviewers used the clinical labels ADD or ADHD to explain why tools had been helpful to them (*"As someone with ADHD, the pomodoro technique has been super effective"*, R615, Forest \typeTrack\typeReward \textbf{P}).

@@

DSCTs needed to accurately capture users' personal definitions of 'distracting use' to be helpful (n = 237, 16\%), a theme often expressed in the form of feature suggestions that would make tools more customisable.

One aspect involved getting tools to accurately capture \emph{what} was considered a distraction.
Thus, for block/removal tools, reviewers wanted to be able to customise which apps or websites were targeted, or target blocking to specific functionality (\emph{"I would like to block reddit but still be able to visit productivity related subreddits"}, R1401, Timewarp \typeBlock\typeTrack\typeGoal \textbf{C}).
Similarly, self-tracking tools often received feature requests for the ability to exempt specific functionality from tracking because not all use represented use that reviewers wanted support in managing (*"I use a meditation app and can be meditating several times a day, sometimes up to an hour at a time, so as it doesn't distinguish between that and social media it gives a false impression of"usage""*, R989, Mute Screen Time Tracker \typeGoal\typeTrack \textbf{A}).

Another aspect involved \emph{when} specific activities should count as distractions.
For example, tools for curbing nightly use received feature requests to handle variation in bedtime schedules (*"it needs to allow for the real life fact that people can't go to bed and wake up at the same time everyday. If it was done on your amount of sleep you get in hours that would be better"*, R1250, [SleepTown](https://apps.apple.com/gb/app/sleeptown/id1210251567)\typeTrack\typeReward \textbf{A}).

**Sub-theme: Alignment between users' and tools' framing of distraction as important for navigating negative emotions elicited by DSCTs**
Aside from practical reasons that DSCTs should accurately capture behaviour that reviewers wished to change, an important factor was the negative emotions that interventions could elicit.
Thus, whereas the various restrictions and reminders provided by DSCTs were seen as helpful and essential, users also needed to carefully navigate the emotional reactions they could create (*"I hate it and love it at the same time. \emojiTearsOfJoy"*, R1299, Stay Focused \typeBlock\typeTrack \textbf{P}, *"Pisses me off sometimes, but worth it"*, R357, Detox Procrastination Blocker, \typeBlock\typeTrack \textbf{P}). 
In order for experiences of "shock", "pressure", etc., to be useful, it was important that the user felt that a DSCT's purpose and framing of usage were closely aligned with their personal perspective.

Thus, tools that provided standards for 'good' or 'bad' usage, for example in their presentation of usage statistics or reminder pop-ups, needed to tread carefully.
Being reminded that one did not behave in line with one's intentions could be very helpful (*"This tells it like I need to be told sometimes...LOVE IT! Simple and effective :)"*, R690, \href{https://chrome.google.com/webstore/detail/go-fucking-work/hibmkkpfegfiinilnlabbfnjcopdiiig}{Go Fucking Work} \typeBlock\typeReward \textbf{C}; *"I like the guilt trip it gives me if I attempt to get into facebook or other social networks"*, R311, Dayboard, \typeBlock\typeGoal \textbf{C}). 
However, normative messaging could also backfire and be profoundly annoying if reviewers did not agree with it (*"I was a bit annoyed when Big Brother started getting judgemental on me. “You have hit your target of 50 pickups!” “You’ve now picked up your phone 90 times!” Well, first I didn’t set a target, and second, I don’t think that 54 minutes across a day should make me feel guilty. Especially since I was reading a book during that time"* R911, Moment \typeGoal\typeTrack \textbf{A}).

@@

Reviewers searched for DSCTs that provided interventions whose incentives were 'just right' to bring about an intended change in behaviour without feeling overly intrusive.
The reviews implied a spectrum of individual differences in where this 'Goldilocks level' was. A few reviewers stated that simply receiving reminders or knowing that a tracking tool was monitoring them was sufficient (*"I know it's tracking me so I pick up the phone a lot less"*, R992, \emph{Mute Screen Time Tracker} \typeGoal\typeTrack \textbf{A}). More commonly, however, reviewers wanted DSCTs to provide more strict or persuasive ways to hold themselves accountable, so that they did not succumb to temptations or unwanted habits in moments of weakness or absentmindedness (n = 132, 9\%).

Thus, tools implementing block/removal design patterns were often criticised if they were too easy to disable (*"This is fantastic but the thing is I can simply remove the extension from chrome which defeats the purpose. Does anyone know how I can stop myself from removing the extension?"*, R1217, Simple Blocker \typeBlock \textbf{C}).
However, if they were *too* difficult to override, they ran into another problem, namely how to deal with emergencies or otherwise unforeseen circumstances.
One developer followed reviewers' suggestion to add a 'pause' button, only to receive critique that blocking was now too easy to override.
Reviewers suggested that adding varying forms of friction to turn off blocking would be a solution, such as having to complete a time-consuming or effortful task (*"make it necessary to touch the screen for a minute or 2 so that it becomes harder for anyone to stop it"*, R8, ( OffTime ) \typeBlock\typeTrack \textbf{P}), or simply adding passwords, which could be set by a trusted other.

Similarly, for tools implementing reward/punishment, some reviewers suggested increasing incentives to make tools more effective (*"If I could also name the trees (so I'll feel even worse about letting them die) it would be even better!"*, R650, Forest \typeTrack\typeReward \textbf{C}). Here too, however, tools needed to find the right balance, because punishments that were perceived as too harsh might make users feel like failures (*"I hate the fact that we get a destroyed building (...) I feel like it is too punishing and almost says 'you've failed'"*, R1247, \href{https://itunes.apple.com/gb/app/sleeptown/id1210251567?mt=8\&uo=4}{SleepTown} \typeTrack\typeReward \textbf{P}).

For tools including goal advancement design patterns, the 'right' level of incentive often related to the frequency or framing of features such as reminder pop-ups, which needed to be finely calibrated to be helpful rather than annoying (*"the todo popup on each page load is quite useful (...) it should be limited (across multiple tabs) to only pop-up every minute or so (...) I find myself snoozing it too often just because the popup bothers me"*, R339, Dayboard, \typeBlock\typeGoal \textbf{C}).

@@

Similarly, some reviewers wanted block/removal tools to not simply block distractions, but also actively nudge their behaviour in ways that supported intended habits.
This might happen by, for example, forwarding blocked websites to productive alternatives, or by replacing algorithmic recommendations with content that they wanted to engage with more ("\emph{if, where the home-screen recommended section is, this could be replaced by a website of our choice} (\ldots) \emph{This way I could do music theory, or learn German. Things I always intend to do, but don't do enough. But if these things came up every time I log into Youtube I'd definitely use them."}, R372, \href{https://chrome.google.com/webstore/detail/df-tube-distraction-free/mjdepdfccjgcndkmemponafgioodelna?hl=en}{DF Tube} \typeBlock \textbf{C}).

>>>>

In academia, a growing amount of HCI research is focusing on developing and evaluating various design patterns meant to help users self-regulate digital device use [@lottridgeBrowserDesignImpacts2012; @kimGoalKeeperExploringInteraction2019; @pinderPushAwaySmartphone2019; @tsengOvercomingDistractionsTransitions2019; @mongeroffarelloRaceDigitalWellbeing2019; @lyngsSelfControlCyberspaceApplying2019][^1].
This area of HCI work has gradually been building an understanding of how a range of design patterns from visualisations of time spent [@whittakerDonWasteMy2016] to goal-setting with social support [@koNUGUGroupbasedIntervention2015] can help people stay in control. However, we are still in the early days of answering basic questions including how to define and measure 'success' for these design patterns, how much people vary in how useful they find particular interventions, and how specific design patterns' usefulness depends on details of their implementation, such as the level of 'friction' they introduce [@kimGoalKeeperExploringInteraction2019; @markEffectsIndividualDifferences2018; @tseng2019; @lyngsTellMeWhat2018; @hinikerMovingGreatScreen2019].

[^1]: We use the term 'design patterns' to refer to specific functions intended to support people in self-regulating their digital device use (e.g., hiding newsfeeds, providing goal reminders), and the term 'digital self-control tool' (DSCT) to refer to apps and browser extensions that implement a given design pattern (e.g., a browser extension for hiding Facebook's newsfeed).
    We occasionally use the term 'intervention' to refer to the implementation of a design pattern in the context of it changing a user's usual digital environment, or when it is evaluated in a research study.

@@

Users want DSCTs to flexibly adapt to their personal definition of 'distraction', as seen in common requests for customisation of *what* and *when* particular usage patterns should be blocked or tracked.
A key reason that users' and tools' conceptualisation of distraction should be in close alignment is that DSCTs' interventions, useful as they may be, can elicit negative emotional reactions.
This may cause the effects of interventions to be counterproductive, especially when users perceive DSCTs' goals or framing to differ from their own.

Finally, users want DSCTs to provide carefully calibrated incentives that are sufficient to incentivise intended behaviour change without feeling overly intrusive.
In our analysis, this 'Goldilocks level'[^2] varied between reviewers, but they commonly wanted tools to provide incentives powerful enough to keep themselves on track in moments of temptation or absentmindedness.
Overly strong incentives could backfire if they simply made users feel like a failure (in the case of reward/punishment tools) or did not allow some way to handle genuine emergency use (in the case of block/removal tools).

[^2]: After the fairy tale 'Goldilocks and the Three Bears', where a young girl tastes three different bowls of porridge and prefers the one that is neither too hot nor too cold, but 'just right'.
    The concept of the 'just right amount' is often referred to as the 'Goldilocks principle'.
    For example, a widely cited paper on the relationship between screen time and mental well-being documented that a balance point in between "too much" and "too little" tech use seemed optimal, which they referred to as the 'digital Goldilocks hypothesis' [@Przybylski2017].

@@

One approach to addressing these questions is to evaluate new or existing interventions in lab and field studies, and/or conduct surveys or interviews with users.
A complementary approach, which we explore in the present paper, is to investigate the landscape of digital self-control tools available online.
Here, many hundreds of tools are exploring a wide range of strategies, from limiting the amount of functionality available on home screens (e.g.\~the Android app \emph{LessPhone Launcher} [@mohanLessPhoneLauncher2019]) and hiding or modifying content on distracting websites (e.g. *Newsfeed Eradicator* [@jdevNewsFeedEradicator2019]), to gamifying self-control by tying device use to the wellbeing of virtual creatures (e.g. *Forest* [@seekrtechForestStayFocused2018]).

@@

We investigated the 367 digital self-control tools reviewed by Lyngs et al. [@lyngsSelfControlCyberspaceApplying2019], as this paper provided the largest currently available review of DSCTs.
Lyngs et al. identified these tools by downloading search results for the terms 'distraction', 'addiction', 'motivation', and 'self-control', individually as well as prefixed by 'internet', 'smartphone', and 'laptop', on the Chrome Web, Google Play, and Apple App stores.
From the initial results (4890 distinct apps and extensions), they included tools explicitly designed to help people self-regulate digital device use, resulting in a data set of 367 tools (222 browser extensions from Chrome Web, 87 apps from Google Play, and 58 apps from the Apple App store).[^3]

[^3]: See Lyngs et al. [@lyngsSelfControlCyberspaceApplying2019] and their supplementary materials on [osf.io/zyj4h/](https://osf.io/zyj4h/) for further details.

@@

## Data and analysis approach

### User numbers and ratings

All stores displayed average ratings, as well as a count of how many ratings a tool had received.
For installation numbers, the stores provided different information: the Chrome Web store provided an exact number of users, and the Google Play store provided a 'minimum number of installs' (e.g. '100,000+').
The Apple App store provided no direct information about user numbers.
For processing which ranked tools by installation numbers, we used number of ratings as a proxy for relative numbers of users for tools on the Apple App store.

@@

-   \typeBlock\hspace{1mm}= block/removal, \typeTrack\hspace{1mm}= self-tracking,\
    \typeGoal\hspace{1mm}= goal-advancement, \typeReward\hspace{1mm}= reward/punish
-   C = Chrome Web, P = Google Play, A = Apple App store
-   Example review quote from a Chrome Web tool including design patterns for block/removal, self-tracking, and goal-advancement: *"It would be nice to make it hard to disable, something like a simple maths problem that we could change difficulty, so I am not tempted."*, R1408, Timewarp \typeBlock\typeTrack\typeGoal \\textbf{C}

@@

An early observation as we explored reported benefits of use was that a substantial proportion of reviews expressed highly positive and important impact of the tools used (n = 480, 22% of reviews).
Thus, the reviews described a common feeling of lacking control over digital device use, which resulted in using devices in unintended ways, or in using them so much that it interfered with other goals.
However, digital self-control tools were seen as effectively helping reviewers take back control, save time, and improve their lives (*"I can happily say that this extension has changed my life!!!! It hides EVERY ANNOYING YouTube feature"*, R385, \href{https://chrome.google.com/webstore/detail/df-tube-distraction-free/mjdepdfccjgcndkmemponafgioodelna?hl=en}{DF Tube} \typeBlock \\textbf{C}, *"Has saved me countless hours of time-wasting nonsense"*, R1000, NewsFeed Eradicator for Facebook \typeBlock\typeGoal \\textbf{C}), and many expressed deep appreciation towards the developers ("Who so ever is working behind this software, my deepest gratitude to them", R271, \href{https://chrome.google.com/webstore/detail/blocksite-block-websites/eiimnmioipafcokbfikbljfdeojpcgbh?hl=en-GB}{\emph{BlockSite}} \typeBlock \\textbf{C}).
Some reviewers seemed surprised that particular tools had been useful to them (*"I added it more as a joke, but it's really good and helpful for those hours you're supposed to study and don't want to be distracted"*, R675, \href{https://chrome.google.com/webstore/detail/go-fucking-work/hibmkkpfegfiinilnlabbfnjcopdiiig}{Go Fucking Work} \typeBlock\typeReward \\textbf{C}).
Importantly, the reviews also contained a wealth of information on how specific implementations of design patterns could fail to fulfil their purpose, and how they might be improved.

@@

Users saw DSCTs as critical to avoiding time-wasting patterns of habitual or impulsive digital device use.
Many reviews described this challenge as related to 'addictive' design patterns and other adversarial dynamics online, to which DSCTs provided essential countermeasures that reviewers had searched for (*"Was looking for a solution for this for over 1 year"*, R771, \href{https://chrome.google.com/webstore/detail/inbox-when-ready-for-gmai/cdedhgmbfjhobfnphaoihdfmnjidcpim}{Inbox When Ready for Gmail} \typeBlock\typeGoal\typeTrack \\textbf{C}; *"Should be on all phones as standard"*, R982, Mute Screen Time Tracker \typeGoal\typeTrack \\textbf{A}).
DSCTs seemed especially helpful when it was not possible to avoid using distracting services (*"This helps me avoid the mind numbing news feed that is of Facebook when work has to be done on Facebook itself"*, R873, \href{https://chrome.google.com/webstore/detail/kill-news-feed/hjobfcedfgohjkaieocljfcppjbkglfd}{Kill News Feed} \typeBlock \\textbf{C}).

The usage contexts mentioned suggested that, aside from wanting to limit time on devices overall or on specific services (n = 62, 4% of reviews), reviewers used DSTCs to help them focus on a given task amidst all too easily available distractions on their devices (n = 201, 13%).
The need for support from DSCTs to stay focused seemed especially acute when tasks were important, but less instantly rewarding, expressed in some reviews as DSCTs helping overcome 'procrastination' (*"seeing my short-term to-do list nudges me to get back to work when I am opening a new website to procrastinate"*, R324, Dayboard \typeBlock\typeTrack \\textbf{C}).
In these situations, DSCTs helped reviewers stay in control by reminding them of what they wanted to do by reducing exposure to 'temptations' (*"Without it, it is too easy for me to be tempted to view and open YouTube sidebar links"*, R376, DF Tube \typeBlock \\textbf{C}), and/or by providing incentives for intended use (*"Not only does it help me stay focused, I smile with delight looking at all my trees and flowers at the end of the day"*, R642, Forest \typeTrack\typeReward \\textbf{C}).
Thus, many reviewers generically stated that DSCTs benefited them in productivity/work contexts (n = 167, 11%; "Not being distracted by unread/misc emails has proven to be of significant benefit to my productivity", R776, \href{https://chrome.google.com/webstore/detail/inbox-when-ready-for-gmai/cdedhgmbfjhobfnphaoihdfmnjidcpim}{\emph{Inbox When Ready for Gmail}} \typeBlock\typeGoal\typeTrack \\textbf{C}), including studying, writing, or reading (n = 90, 6%).

Another important use context was \emph{sleep}, where many reviewers said their digital devices often interfered with going to bed at the intended time.
However, DSCTs helped them adjust their habits in favour of the sleep patterns they wanted (*"I always used to stay up really late on my phone until I got this app - I have consistently been getting 9 hours sleep and waking up early for the first time in my life !!"*, R1252, [SleepTown](https://apps.apple.com/gb/app/sleeptown/id1210251567)\typeTrack\typeReward \\textbf{P}; *"It really forces me to stay off to the phone, and so it's useful especially when I need to go to bed"*, R198, BlackOut \typeBlock \\textbf{P}).

**Sub-theme: DSCTs as especially useful for people with more acute self-regulation challenges** In an extension of the main theme, DSCTs seemed especially useful for individuals who experience self-regulation challenges more intensely.
This was often expressed as a tool being helpful for tackling 'addiction', and/or as reviewers labelling themselves 'addicts', 'procrastinators' or just low in self-control (n = 58, 3.8%; *"This was the perfect solution to my phone addiction. It was consuming my life and my free time"*, R612, Forest \typeTrack\typeReward \\textbf{P}).
Sometimes reviewers used the clinical labels ADD or ADHD to explain why tools had been helpful to them (*"As someone with ADHD, the pomodoro technique has been super effective"*, R615, Forest \typeTrack\typeReward \\textbf{P}).

@@

DSCTs needed to accurately capture users' personal definitions of 'distracting use' to be helpful (n = 237, 16%), a theme often expressed in the form of feature suggestions that would make tools more customisable.

One aspect involved getting tools to accurately capture \emph{what} was considered a distraction.
Thus, for block/removal tools, reviewers wanted to be able to customise which apps or websites were targeted, or target blocking to specific functionality (\emph{"I would like to block reddit but still be able to visit productivity related subreddits"}, R1401, Timewarp \typeBlock\typeTrack\typeGoal \\textbf{C}).
Similarly, self-tracking tools often received feature requests for the ability to exempt specific functionality from tracking because not all use represented use that reviewers wanted support in managing (*"I use a meditation app and can be meditating several times a day, sometimes up to an hour at a time, so as it doesn't distinguish between that and social media it gives a false impression of"usage""*, R989, Mute Screen Time Tracker \typeGoal\typeTrack \\textbf{A}).

Another aspect involved \emph{when} specific activities should count as distractions.
For example, tools for curbing nightly use received feature requests to handle variation in bedtime schedules (*"it needs to allow for the real life fact that people can't go to bed and wake up at the same time everyday. If it was done on your amount of sleep you get in hours that would be better"*, R1250, [SleepTown](https://apps.apple.com/gb/app/sleeptown/id1210251567)\typeTrack\typeReward \\textbf{A}).

**Sub-theme: Alignment between users' and tools' framing of distraction as important for navigating negative emotions elicited by DSCTs** Aside from practical reasons that DSCTs should accurately capture behaviour that reviewers wished to change, an important factor was the negative emotions that interventions could elicit.
Thus, whereas the various restrictions and reminders provided by DSCTs were seen as helpful and essential, users also needed to carefully navigate the emotional reactions they could create (*"I hate it and love it at the same time.* \emojiTearsOfJoy", R1299, Stay Focused \typeBlock\typeTrack \\textbf{P}, *"Pisses me off sometimes, but worth it"*, R357, Detox Procrastination Blocker, \typeBlock\typeTrack \\textbf{P}).
In order for experiences of "shock", "pressure", etc., to be useful, it was important that the user felt that a DSCT's purpose and framing of usage were closely aligned with their personal perspective.

Thus, tools that provided standards for 'good' or 'bad' usage, for example in their presentation of usage statistics or reminder pop-ups, needed to tread carefully.
Being reminded that one did not behave in line with one's intentions could be very helpful (*"This tells it like I need to be told sometimes...LOVE IT! Simple and effective :)"*, R690, \href{https://chrome.google.com/webstore/detail/go-fucking-work/hibmkkpfegfiinilnlabbfnjcopdiiig}{Go Fucking Work} \typeBlock\typeReward \\textbf{C}; *"I like the guilt trip it gives me if I attempt to get into facebook or other social networks"*, R311, Dayboard, \typeBlock\typeGoal \\textbf{C}). However, normative messaging could also backfire and be profoundly annoying if reviewers did not agree with it (*"I was a bit annoyed when Big Brother started getting judgemental on me. "You have hit your target of 50 pickups!" "You've now picked up your phone 90 times!" Well, first I didn't set a target, and second, I don't think that 54 minutes across a day should make me feel guilty. Especially since I was reading a book during that time"* R911, Moment \typeGoal\typeTrack \\textbf{A}).

@@

Reviewers searched for DSCTs that provided interventions whose incentives were 'just right' to bring about an intended change in behaviour without feeling overly intrusive.
The reviews implied a spectrum of individual differences in where this 'Goldilocks level' was.
A few reviewers stated that simply receiving reminders or knowing that a tracking tool was monitoring them was sufficient (*"I know it's tracking me so I pick up the phone a lot less"*, R992, \emph{Mute Screen Time Tracker} \typeGoal\typeTrack \\textbf{A}).
More commonly, however, reviewers wanted DSCTs to provide more strict or persuasive ways to hold themselves accountable, so that they did not succumb to temptations or unwanted habits in moments of weakness or absentmindedness (n = 132, 9%).

Thus, tools implementing block/removal design patterns were often criticised if they were too easy to disable (*"This is fantastic but the thing is I can simply remove the extension from chrome which defeats the purpose. Does anyone know how I can stop myself from removing the extension?"*, R1217, Simple Blocker \typeBlock \\textbf{C}).
However, if they were *too* difficult to override, they ran into another problem, namely how to deal with emergencies or otherwise unforeseen circumstances.
One developer followed reviewers' suggestion to add a 'pause' button, only to receive critique that blocking was now too easy to override.
Reviewers suggested that adding varying forms of friction to turn off blocking would be a solution, such as having to complete a time-consuming or effortful task (*"make it necessary to touch the screen for a minute or 2 so that it becomes harder for anyone to stop it"*, R8, ( OffTime ) \typeBlock\typeTrack \\textbf{P}), or simply adding passwords, which could be set by a trusted other.

Similarly, for tools implementing reward/punishment, some reviewers suggested increasing incentives to make tools more effective (*"If I could also name the trees (so I'll feel even worse about letting them die) it would be even better!"*, R650, Forest \typeTrack\typeReward \\textbf{C}). Here too, however, tools needed to find the right balance, because punishments that were perceived as too harsh might make users feel like failures (*"I hate the fact that we get a destroyed building (...) I feel like it is too punishing and almost says 'you've failed'"*, R1247, \href{https://itunes.apple.com/gb/app/sleeptown/id1210251567?mt=8\&uo=4}{SleepTown} \typeTrack\typeReward \\textbf{P}).

For tools including goal advancement design patterns, the 'right' level of incentive often related to the frequency or framing of features such as reminder pop-ups, which needed to be finely calibrated to be helpful rather than annoying (*"the todo popup on each page load is quite useful (...) it should be limited (across multiple tabs) to only pop-up every minute or so (...) I find myself snoozing it too often just because the popup bothers me"*, R339, Dayboard, \typeBlock\typeGoal \\textbf{C}).

@@

Similarly, some reviewers wanted block/removal tools to not simply block distractions, but also actively nudge their behaviour in ways that supported intended habits.
This might happen by, for example, forwarding blocked websites to productive alternatives, or by replacing algorithmic recommendations with content that they wanted to engage with more ("\emph{if, where the home-screen recommended section is, this could be replaced by a website of our choice} (\ldots) \emph{This way I could do music theory, or learn German. Things I always intend to do, but don't do enough. But if these things came up every time I log into Youtube I'd definitely use them."}, R372, \href{https://chrome.google.com/webstore/detail/df-tube-distraction-free/mjdepdfccjgcndkmemponafgioodelna?hl=en}{DF Tube} \typeBlock \\textbf{C}).

**Sub-theme 2: DSCTs should have low friction to start and use, but high friction to stop** Finally, finding the right level of friction also related to the level of effort required to use a tool.
The amount of effort needed to \emph{initiate} use was important to keep low, because a lack of motivation could easily lead users to not use an otherwise effective tool (*"Most users have procrastination issues which make them postpone the moment in which they will click on the "Start focusing" button. Therefore those users (including me) will get distracted and they will click on that button after several hours, or even never, sometimes wasting a whole working day"*, R545, Focus \typeBlock\typeGoal\typeTrack \\textbf{C}). Therefore, a common feature request for blocking tools was the ability to set a schedule where blocking would automatically turn on. An interesting suggestion related to blocking tools alternating between work and break cycles: these could flip the ordering of cycles so as to begin with a break, followed by enforced blocking, which might lower the amount of motivation needed to initiate a focus session (*"Hitting the start button the very first time is hard! Starting in a short break would allow me to hit this without a second thought, because I have a few short moments to surf before the focus is forced."*, R232, Block & Focus \typeBlock\typeTrack \\textbf{C}).
